Parsing Postfix log files

### 2008-11-06 doc/logparser-background.tex
Dealt with some XXX notes.

Each restriction is evaluated to produce a result of \textit{reject},
\textit{permit}, \textit{dunno\/} or the name of another restriction to be
evaluated.\footnote{Other results are possible as described
in~\cite{smtpd_access_readme,smtpd_per_user_control,policy-servers}.} The
meaning of \textit{permit\/} and \textit{reject\/} is fairly obvious;
\textit{dunno\/} means to stop evaluating the current restriction and
continue processing the remainder of the restriction list, allowing
exceptions to more general rules.  When the result is the name of another
restriction Postfix will evaluate that restriction, allowing restrictions
to be chosen based on the client \IP{} address, client hostname, HELO
hostname, sender address, or recipient address.\footnote{E.g.\ the
administrator may require that all clients on the local network have valid
DNS entries, to prevent people sending mail from unknown machines.}  The
administrator can define new restrictions as a list of existing
restrictions, allowing arbitrarily long and complex sequences of lookups,
restrictions and exceptions.  Postfix tries to protect the administrator in
as far as is reasonable, e.g.\ the restriction
\texttt{check\_helo\_mx\_access} cannot cause a mail to be accepted,
because the parameter it checks (the hostname given in the HELO command) is
under the control of the remote client.  Despite this, it is possible for
the administrator to make catastrophic mistakes, e.g.\ rejecting all mail
--- the administrator must be cognisant of the effects of their
configuration changes.  This is similar to one of UNIX's design
philosophies: \textit{``UNIX was not designed to stop its users from doing
stupid things, as that would also stop them from doing clever
things''\/}~\cite{unix-philosophy}.

@@

The restriction \texttt{check\_client\_access} checks if the \IP{} address
of the connected client is found in the specified file and returns the
associated action if found; the method of searching the file is dependant
on the type of the file (\texttt{cidr} in the example) --- see
\cite{postfix-lookup-tables} for more details.  Other restrictions
determine their result by consulting external sources, e.g.\newline{}
\tab{}\texttt{reject\_rbl\_client dnsbl.example.com}\newline{} checks the
\DNSBL{} \texttt{dnsbl.example.com}, rejecting the command if the client's
\IP{} address is listed.  XXX DO I NEED THIS SENTENCE\@?  This description
is necessarily brief, for further details see~\cite{smtpd_access_readme,
smtpd_per_user_control, policy-servers}.

@@

    XXX HOW CAN I CENTRE THE TABLE\@?\hfill{}

>>>>

Each restriction is evaluated to produce a result of \textit{reject},
\textit{permit}, \textit{dunno\/} or the name of another restriction to be
evaluated.\footnote{Other results are possible as described
in~\cite{smtpd_access_readme,smtpd_per_user_control,policy-servers}.} The
meaning of \textit{permit\/} and \textit{reject\/} is fairly obvious;
\textit{dunno\/} means to stop evaluating the current restriction and
continue processing the remainder of the restriction list, allowing
exceptions to more general rules.  When the result is the name of another
restriction Postfix will evaluate that restriction, allowing restrictions
to be chosen based on the client \IP{} address, client hostname, HELO
hostname, sender address, or recipient address.\footnote{E.g.\ the
administrator may require that all clients on the local network have valid
DNS entries, to prevent people sending mail from unknown machines.}  The
administrator can define new restrictions as a list of existing
restrictions, allowing arbitrarily long and complex sequences of lookups,
restrictions and exceptions.  Postfix tries to protect the administrator in
as far as is reasonable, e.g.\ the restriction
\texttt{check\_helo\_mx\_access} cannot cause a mail to be accepted,
because the parameter it checks (the hostname given in the HELO command) is
under the control of the remote client.  Despite this, it is possible for
the administrator to make catastrophic mistakes, e.g.\ rejecting all mail
--- the administrator must be cognisant of the effects their configuration
changes will have.  This is similar to one of UNIX's design philosophies:
\textit{``UNIX was not designed to stop its users from doing stupid things,
as that would also stop them from doing clever
things''\/}~\cite{unix-philosophy}.

@@

The restriction \texttt{check\_client\_access} checks if the \IP{} address
of the connected client is found in the specified table and returns the
associated action if found; the method of searching the file is dependant
on the type of the file (\texttt{cidr} in the example) --- see
\cite{postfix-lookup-tables} for more details.  Other restrictions
determine their result by consulting external sources, e.g.\newline{}
\tab{}\texttt{reject\_rbl\_client dnsbl.example.com}\newline{} checks the
\DNSBL{} \texttt{dnsbl.example.com} and rejects the command if the client's
\IP{} address is listed.

For further information about Postfix restrictions
see~\cite{smtpd_access_readme,smtpd_per_user_control,policy-servers}.

@@

    \centering{}

### 2008-11-06 doc/logparser-conclusion.tex
Dealt with some XXX notes.

The division of the parser into rules, actions, and framework is unusual
because rules are separated so completely from the actions and framework;
although parsers are often divided (the combination of \texttt{lex} and
\texttt{yacc} XXX BOOK REF FOR LEX AND YACC being a common example), the
parts are usually quite internally interdependent, and will be combined
into a complete parser by the compilation process; in contrast
\parsername{} keeps the rules and actions separate until the parser runs.
The actions and framework are not as completely separated, as the actions
depend on services provided by the framework; however the actions and
framework are not tightly integrated: it would be possible, with some work,
to completely separate the two.  XXX I SHOULD DO THAT\@.

>>>>

The division of the parser into rules, actions, and framework is unusual
because rules are separated so completely from the actions and framework;
although parsers are often divided (the combination of \texttt{lex} and
\texttt{yacc}~\cite{lex-and-yacc-book} being a common example), the parts
are usually quite internally interdependent, and will be combined into a
complete parser by the compilation process; in contrast \parsername{} keeps
the rules and actions separate until the parser runs.  The actions and
framework are not as completely separated, as the actions depend on
services provided by the framework; however the actions and framework are
not tightly integrated: it would be possible, with some work, to completely
separate the two.  XXX I SHOULD DO THAT\@.

### 2008-11-06 doc/logparser-implementation.tex
Rearranged sections; no real content changes.
Dealt with some XXX notes.

\section{Postfix Parser Implementation}

@@

XXX SHOULD I USE TEXTTT OR SIMILAR ON FIELD NAMES\@?

@@

    \item [postfix\_action] This is the action Postfix must have taken to
        generate this log line.  This field is mostly ignored, but two
        values (IGNORED and INFO) have special meaning, as described below
        in the list of typical values.\label{postfix_action}

        XXX MAYBE MOVE THIS LIST TO AFTER THE FIELD LIST\@?  DO I NEED THIS
        LIST AT ALL\@?

@@

        Uninteresting log lines are parsed so that any lines the parser
        is not capable of handling become immediately obvious errors.

@@

@@

XXX MOVE TO AFTER ACTIONS\@.

@@

    \item [MAIL\_QUEUED] This action represents Postfix picking a mail from
        the queue to deliver. This action is used for both \daemon{qmgr}
        and \daemon{cleanup} as it needs to deal with out of order log
        lines; see \sectionref{discarding cleanup lines} for details.  XXX
        EXPLAIN THIS BETTER\@.

>>>>

\section{Postfix Parser Implementation}
\label{Postfix Parser Implementation}

@@

@@

    \item [postfix\_action] The action Postfix must have taken to generate
        this log line; the possible actions are described later in
        \sectionref{list of postfix actions}.

\label{list of postfix actions}

@@

@@

    \item [MAIL\_QUEUED] This action represents Postfix picking a mail from
        the queue to deliver. This action is used for both \daemon{qmgr}
        and \daemon{cleanup} as it needs to deal with out of order log
        lines; see \sectionref{discarding cleanup lines} for details.
        XXXXXX EXPLAIN THIS BETTER\@.

\subsection{Framework}

XXX TO BE WRITTEN\@.

@@

@@

### 2008-11-06 doc/logparser-introduction.tex
Dealt with some XXX notes.

Section~XXX contains the core of the paper, describing
a naive parsing algorithm and the complications initially encountered that
shaped the full algorithm.  A flow chart and a discussion of the emergent
behaviour exhibited by the algorithm accompanies a comprehensive
explanation of the different stages of the initial algorithm.  The
framework that actions and rules fit into is documented, then the actions
taken during execution of the algorithm are described, followed by the
process of adding a new action.  The section concludes with an in-depth
description of the further complications discovered, and their solutions
that complete the parser.

>>>>

Section~\ref{Postfix Parser Implementation} contains the core of the paper,
describing a naive parsing algorithm and the complications initially
encountered that shaped the full algorithm.  A flow chart and a discussion
of the emergent behaviour exhibited by the algorithm accompanies a
comprehensive explanation of the different stages of the initial algorithm.
The framework that actions and rules fit into is documented, then the
actions taken during execution of the algorithm are described, followed by
the process of adding a new action.  The section concludes with an in-depth
description of the further complications discovered, and their solutions
that complete the parser.

### 2008-11-06 doc/logparser-other-parsers-review.tex
Dealt with some XXX notes.

There only appears to be one prior published paper about parsing Postfix
log files: \textit{Log Mail Analyzer: Architecture and Practical
Utilizations\/}~\cite{log-mail-analyser}.  The aim of \LMA{} is quite
different from this parser: it attempts to present correlated log data in a
form suitable for a systems administrator to search using the myriad of
standard Unix text processing utilities already available, producing both a
\CSV{} file and either a MySQL (\url{http://www.mysql.com/}) or Berkeley DB
(\url{http://www.oracle.com/database/berkeley-db/index.html}) database.
XXX SHOULD THOSE URLS BE REFERENCES INSTEAD\@?  The decision to support
both \CSV{} and Berkeley DB appears to have been a serious limitation:
\CSV{} is a very simple format where a record is stored in a single line,
with fields separated by a comma or other punctuation symbol.  Problems
with \CSV{} files include the need to escape separators in the data stored,
providing multiple values for a field (e.g.\ multiple recipients), and
adding new fields.  There is no mechanism to document the fields or the
separator, unlike \SQL{} databases where every database includes a schema
naming the fields and the type of data they store (integer, text,
timestamp, etc.).

@@

Isoqlog's report lists daily, monthly, and annual totals of mails sent by
each unique sender and bytes transferred (XXX EXPLAIN BYTES TRANSFERRED\@:
TO/FROM, PER-SENDER\@?).  It only produces these reports for the domains
listed in its configuration file, making it impossible to produce reports
for every sender domain.  It ignores all log lines except those for today,
so it is impossible to analyse historical logs, and testing with the
\numberOFlogFILES{} test log files is pointless.  It does maintain a record
of data previously extracted, which the newly extracted data is merged into
(no information is provided on the format of the data store).  The data
extracted is limited to the number of mails sent by each sender, unlike the
copious amounts of data extracted by \parsername{}.  It does not utilise
rejection log lines in any way, so is unsuitable for the purposes of this
project.  Its parsing is completely inextensible, indeed is almost
incomprehensible, relying on \texttt{scanf(3)}, unexplained fixed offsets,
and low level string manipulation; it is the opposite end of the spectrum
to \parsernames{} parsing.  It does not handle any of the complications
discussed in \sectionref{complications}, does not gather the breadth of
data required for this project, and ignores the majority of log lines
produced by Postfix.

>>>>

There only appears to be one prior published paper about parsing Postfix
log files: \textit{Log Mail Analyzer: Architecture and Practical
Utilizations\/}~\cite{log-mail-analyser}.  The aim of \LMA{} is quite
different from this parser: it attempts to present correlated log data in a
form suitable for a systems administrator to search using the myriad of
standard Unix text processing utilities already available, producing both a
\CSV{} file and either a MySQL (\url{http://www.mysql.com/}) or Berkeley DB
(\url{http://www.oracle.com/database/berkeley-db/index.html}) database.
The decision to support both \CSV{} and Berkeley DB appears to have been a
serious limitation: \CSV{} is a very simple format where a record is stored
in a single line, with fields separated by a comma or other punctuation
symbol.  Problems with \CSV{} files include the need to escape separators
in the data stored, providing multiple values for a field (e.g.\ multiple
recipients), and adding new fields.  There is no mechanism to document the
fields or the separator, unlike \SQL{} databases where every database
includes a schema naming the fields and the type of data they store
(integer, text, timestamp, etc.).

@@

Isoqlog's report misses most of the information gathered by \parsername{},
and it only reports on the domains listed in its configuration file, making
it impossible to produce reports for every sender domain.  It ignores all
log lines except those for today, so it is impossible to analyse historical
logs, and testing with the \numberOFlogFILES{} test log files is pointless.
It does maintain a record of data previously extracted, which the newly
extracted data is merged into (no information is provided on the format of
the data store).  The data extracted is limited to the number of mails sent
by each sender, unlike the copious amounts of data extracted by
\parsername{}.  It does not utilise rejection log lines in any way, so is
unsuitable for the purposes of this project.  Its parsing is completely
inextensible, indeed is almost incomprehensible, relying on
\texttt{scanf(3)}, unexplained fixed offsets, and low level string
manipulation; it is the opposite end of the spectrum to \parsernames{}
parsing.  It does not handle any of the complications discussed in
\sectionref{complications}, does not gather the breadth of data required
for this project, and ignores the majority of log lines produced by
Postfix.
